### 2024-07-12 rebuttal_07092024.txt
added rebuttal wip

REVIEWER 1)

Please find my comments below:
1. The theory section of the paper is explained in depth and the technical contributions are also articulated in depth.
2. The figures however are very inconsistent. For example, figure 7 uses "Train Grav" in the legend whereas figure 8 uses Gravel. In other cases, asphalt is mentioned as "Asph" or "Asphalt". Please choose one terminology and keep it consistent across all figures.

	The authors made the figures more consistent.

3. The font size in the legends are highly inconsistent. Please set the graphic parameters to be the same.

	The authors made the figures more consistent.

4. The terms "grass" and "unknown/unkn" are used interchangeably. As earlier, it is rather confusing when each plot addresses these categories differently.

	The authors made the terminology more consistent.

5. The line color in Figure 9 is almost the same. Please use contrasting colors to distinguish the estimated min classification error and observed min classification error.

	The authors made the figure more legible.

6. Prior to the results section, the "dirt" category was not mentioned in the training phase. But it seems to appear only in the results section. Could the authors clarify the "dirt" category and how it is introduced in this experiment?

	Dirt was not used in the training phase. Rather, data was simply gathered and used to compare against gravel data. This point was clarified in the text.

7. Some citations are not referenced correctly in the introduction section, for example in the line "Furthermore, RDFs are less complex to train, able to be trained with smaller data sets, and are generally
easier to interpret [? ? ? ? ]."

	Citations have been fixed.

8. The references are not consistent. Reference 7 has the title in capital letters whereas the titles are in sentence case in others. Please keep all of the reference consistent in style.

	References have made more consistent.

Thank you for making the revisions and responding. Many of the reviewer’s previous comments were well-addressed. However, some points in the manuscript still require further clarification.

1) Please do provide the area size of each type of the surfaces in square meters. This provides important fact regarding the data used in this work.

	The authors do not believe that this information is pertinent information to add to the text, as it is not only irrelevant to the work, the potential implications and questions would need to be explained in text. The reasons for this were explained in a previous rebuttal. Unless the reviewer provides a reason as to why this information is valuable, the authors will not be adding this information to the work. As stated in the text: "As the training data is independent of the size of each designated training area and the number of LiDAR points in that area, it is not considered in this work".

2) The reviewer did not find the updates in the manuscript related the following comment from the previous round. Please highlight your changes in the manuscript or copy the changes to the response letter. Ideally both.
Figure 16 (“Side-of-road detection accuracy was…”) and the associated description in the Results section. The features the method used do not sufficiently separate the two classes does not mean the task itself is not possible. It could just be the features we engineered do not have sufficient descriptive power. Having this problem fully solved may be beyond the scope of this work. However, the authors could attempt to provide more insights into the nature of the problem. For example, why is this problem not feasible by its nature? Or, if it is possible, how could it be better solved (with a higher resolution lidar or some specific features, etc.)?

	The authors provided a diff file in the same pdf document as the rebuttal - changes in red are deletions, changes in blue are additions. Meanwhile the submitted document is un-annotated. This practice will continue, however any significant changes will be repeated in this rebuttal. As stated in the document, further distinction between the classes may be accomplished but it outside the scope of this present work. A higher resolution LiDAR may provide more detailed or consistent features with a higher point density.

3) Same as above, the reviewer did not find the updates in the manuscript related the following comment from the previous round. Please highlight your changes in the manuscript or copy the changes to the response letter. Ideally both.
In the Results section, “however, for this study, the algorithm that produced the final results used two thousand points of reference per class.” It is still somewhat confusing regarding the definition of “point.” Does a point mean a lidar point or a time instance that corresponds to, say, five lidar points of each lidar? Could you clarify this further in the manuscript?

	The submitted document does not contain the sentence "however, for this study, the algorithm that produced the final results used two thousand points of reference per class". This was revised as requested in the previous revisions, and the changes were highlighted in the annotated document: "Found research did not make any conclusive arguments for the proper training data set size for Random Decision Forest training, however after thirty percent of the training data was randomly selected and reserved to create a validation data set, approximately two thousand samples of gravel, asphalt, and grass were used to train the algorithm". The authors believe this should alleviate any confusion.

4) Related work: “Lei et al suggest a method using an SVM for the detection of unstructured dirt and gravel roads using a 2D LiDAR and a camera [? ], however the method does not use a 3D scanning LiDAR and cameras may be insufficient for reliable road detection [19]. Fernandez-Delgado et al found that RDFs out-performed other machine learning algorithms, including NNs, when classifying 121 data sets taken from the UCI data base [? ]. Therefore while both Neural Networks and RDFs have been used to classify scanning LiDAR data, a RDF classifier was chosen. Furthermore, RDFs are less complex to train, able to be trained with smaller data sets, and are generally easier to interpret [? ? ? ? ].” Please correct all the [????] in the text. The same problem exists in other sections in the manuscript, e.g., Section Training Data Extraction and Section RDF Training and Testing.

	The authors fixed the citations.

5) Figure 9. Please use different line styles and drastically different colors for the two lines. Also, use a larger marker for the best point. Since the best hyper-parameters are chosen based on the minimum upper confidence interval of the classification error, the confidence interval should be included in the plot as well.

	The authors made the figure more legible.

6) Figure 10. Is the Matlab marker included in the plot by design? If so, please make it more visible. If not, please have it removed.

	Good catch! The authors removed the matlab marker.

7) Table I. Where is this table referenced in the text? Also, add a caption to describe how this table should be read. It is not clear what the three columns of information are currently.

	The table is now referenced in the text. The table was modified to be made more legible.